Powering Hidden Markov Model by Generative Models

### 2019-08-18 manuscript/main.tex
write expmt section

\usepackage{caption}
\usepackage{bm}
\newcommand{\ubar}[1]{\mkern2mu\underline{\mkern-2mu #1\mkern-2mu}\mkern2mu}
% \allowdisplaybreaks
\usepackage{mystyle}
\newcommand{\ubm}[1]{\ubar{\bm{#1}}}
\newcommand{\ubmr}[2]{\ubar{\bm{#1}}^{#2}}

@@

The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: \textcolor{blue}{a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels. Our experiments in this section is carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.

@@

>>>>

\usepackage{bm}
\newcommand{\ubar}[1]{\mkern2mu\underline{\mkern-2mu #1\mkern-2mu}\mkern2mu}
% \allowdisplaybreaks
\usepackage{mystyle}
\newcommand{\ubm}[1]{\ubar{\bm{#1}}}
\newcommand{\ubmr}[2]{\ubar{\bm{#1}}^{#2}}

@@

The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: \textcolor{blue}{a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels. Our experiments in this section is carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}. \textcolor{blue}{The mapping table for the folding process is attached in appendix.}

@@

\begin{table}
  \caption{Test accuracy table for $39$ dimensional features }\label{tab:acc-classification39f}
  \vspace{-0.3cm}
  \centering
  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
  \centering  
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
    GMM-HMM & $62.30\%$ &  $68.06\%$ &  $68.73\%$  \\ \midrule
    GenHMM & $76.78\%$ & $77.68\%$ &   \\ 
    \bottomrule                                                                  
  \end{tabular}
  \bigskip
  \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
  % \centering  
  \begin{tabular}{l|c|c|c} \toprule
    {Model} &  K=1 &  K=3 &  K=5
    \\ \midrule
    GMM-HMM & $53.64\%$ &  $59.65\%$ & $61.98\%$  \\ \midrule
    GenHMM & $69.51\%$ & $70.67\%$ & $70.78\%$   \\ 
    \bottomrule
  \end{tabular}
  \vspace{0.3cm}
\end{table}

We firstly show the phoneme classification using 39 dimensional MFCC features (MFCC coefficients, deltas, and delta-deltas), to validate one possible usage of our proposed model. GMM-HMM is trained and tested as a reference model in our experiments. Training and testing of GMM-HMM is in the same condition as GenHMMs are trained and tested. Dataset usage for GenMM and GMM-HMM is the same, and number of states for GMM-HMM is the same as that for GenHMM in modeling each phoneme. Apart from setting the reference model, we also run the experiment comparisons with different total number of mixture components ($K$).

Table \ref{tab:acc-classification39f} shows the results for this experiments, in which we tested both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and also the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}. As shown in both $61$-phoneme and $39$-phoneme cases, GenHMM gets significant higher accuracy than GMM-HMM for the same number of mixture components. As our expectation, GenHMM should have better modeling capacity of sequential data since we bring in the neural network based generators into GenHMM, which should be able to represent complex relationship between states of HMM and sequential data. This is some increase of accuracy as total number of mixture components in GenHMM is increased from $K=1$ to $K=5$. The sequential dependency of data is modeled by HMM itself, while each state of HMM can have better representation using a mixture probabilistic model if data represented by the state is multi-mode. Comparing the results in $39$-phoneme and $61$-phoneme cases, GenHMM gets higher accuracy for $39$-phoneme classification than it does for $61$-phoneme classification. The total training dataset size remains the same as $61$ phonemes are folded into $39$ phonemes. There are less training data available and more classes to be recognized in the $61$-phoneme case, which makes the take more challenging.

\begin{table}
  \caption{Test accuracy table for $13$ dimensional features }
  \label{tab:acc-classification13f}
  \vspace{-0.3cm}
  \centering
  \subcaption{Folded $39$ Phonemes}
  \centering  
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
    GMM-HMM & $48.59\%$ &  $51.26\%$ &  $52.40\%$  \\ \midrule
    GenHMM & $\%$ & $\%$ &   \\ 
    \bottomrule                                                                  
  \end{tabular}
  \bigskip
  \subcaption{$61$ Phonemes}
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5
    \\ \midrule
    GMM-HMM & $37.15\%$ &  $40.65\%$ & $42.27\%$  \\ \midrule
    GenHMM & $\%$ & $\%$ & $\%$   \\ 
    \bottomrule
  \end{tabular}
  \vspace{0.3cm}
\end{table}

Similar experiments to that in table~\ref{tab:acc-classification13f} are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas). The results are shown in table~\ref{tab:acc-classification13f}. The superior performance of GenHMM remains compared with reference model GMM-HMM. The gain by using mixture generators are also present in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is also similar to the set of experiments in table~\ref{tab:acc-classification39f}.

\section{Conclusion}
